Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-25 paper/agree-semantics.tex
Add minor revisions to section 4

Although basic, these definitions can be used to define higher-level
operators from PTLTL, such as \konst{Once} and \konst{Since}. 
\[
   \begin{array}{rcl}
      \sem{\konst{Once}(b)}^E_t & = & \sem{b \vee (\mathit{false} \to \konst{pre}(\konst{Once}(b)))}^E_t
   \end{array}
\]

>>>>

Although basic, these definitions can be used to define higher-level
past-time operators from PLTL such as \konst{Once} and \konst{Since}.
These are defined relative to their corresponding bound output stream $o$.
\[
   \begin{array}{rcl}
      \sem{\konst{Once}(b)}^E_t & = & \sem{b \vee (\mathit{false} \to \konst{pre}(o))}^E_t \\
      \sem{\konst{Since}(a, b)}^E_t & = & \sem{b \vee (a and (\mathit{false} \to \konst{pre}(o)))}^E_t \\
   \end{array}
\]
$\konst{Once}(b)$ is true if $b$ has ever been true before, or in, the current moment.
$\konst{Since}(a, b)$ is true if $a$ has been true in all moments up to the present one since $b$ most recently became true.
These both refer the their bound output stream $o$ to know if they were true or false in their previous step.
All other past-time operators of PLTL can be defined similarly.

### 2022-03-25 paper/sw-contract.tex
Add minor revisions to section 4

The AGREE specification for the SW component in the example of
Section~\ref{sec:example} is given in \figref{fig:sw}.  The
specification uses \texttt{eq} statements to define variables local to
the contract specification.  For
example, \lineref{line:sw-event-def-start} defines the \texttt{req}
variable to be equivalent to the \texttt{event} expression.

@@

The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract.
Streams are associated with every expression in AGREE, meaning that expressions are defined through time, and are therefore stateful.
When writing assumptions and guarantees, it is important to differentiate pre-state, before a contract updates its state, and post-state, after a contract updates its state, in response to the current input. AGREE provides the \texttt{pre} operator to make that distinction.

The \texttt{pre} operator returns the previous value of the enclosed expression by looking back one step in time on the expression's stream.
In general, assumptions regarding state should be evaluated in the pre-state of the component with the current inputs, and guarantees regarding state should be evaluated on the post-state of the component given the current inputs.
Guarantees should use the \texttt{pre} operator anytime they need to reason about the current state relative to the previous state.

The assumption on \lineref{line:sw-assume-3} uses a \emph{followed-by} operator, \texttt{->}, to \emph{guard} the \texttt{pre} operator at time 0.
A followed-by expression defines what happens at the first instance of the stream, time 0, and then what follows after.
Here the assumption is \texttt{true} at time 0 and then it is the truth value of the implication in all future instances.
The followed-by guards the \texttt{pre} because expressions are undefined before time 0 and the followed-by prevents the \texttt{pre} from being evaluated before time 1.
So after time 0, the assumption depends on the presence of a request and the pre-state of the component.

The stream semantics in AGREE mean that it is possible to use PLTL operators that look back in time.
The \texttt{Historically} operator is the same as $\historically$ defined previously; it is true if the enclosed expression has been true in all moments up to the present one.
The \texttt{Since} operator is true if the first expression has been true in all moments up to the present one since the second expression most recently became true.
The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}--\emph{not request since start}.

The guarantees on \linesref{line:sw-guarantee-1}{line:sw-guarantee-3} coincide events and assert well-formed output.
The guarantees on \linesref{line:sw-guarantee-4}{line:sw-guarantee-6} define temporal properties of the component.
\lineref{line:sw-guarantee-4} insists that a start happens with a request or one step after a request.
The guarantee uses the assumption on \lineref{line:sw-assume-3} and does not check for two requests in a row without a start as the assumption precludes that input behavior.
It differentiates with the followed-by operator what is required at time 0, the start must coincide with the request, with what is required after time 0, the start must coincide with the request or is a response to a request one step earlier.
The guarantee also does not force the start to always happen, it only says that if it does happen, it is under the defined conditions.

\lineref{line:sw-guarantee-5} forces the alert to sound if the start does not arrive within the one-step bound.
Together with \lineref{line:sw-guarantee-4} the contract model allows for non-alerting and alerting behavior.
The last guarantee ensures that alert is persistent.
The \texttt{Once} operator is true if the enclosed expression has ever been true before, or in, the current moment.
\lineref{line:sw-guarantee-6} ensures if alert has ever happened, then it is happening in the present moment.
These six guarantees define the behavior of the SW component under the three assumptions.

>>>>

The actual surface syntax for AGREE that is embedded in OSATE is slightly different than the formal semantics in the previous section.
The mapping between the two is straightforward.
The AGREE specification for the SW component in the example of
Section~\ref{sec:example} is given in \figref{fig:sw}.  The
specification uses \texttt{eq} statements to define variables local to
the contract specification.  For
example, \lineref{line:sw-event-def-start} defines the \texttt{req}
variable to be equivalent to the \texttt{event} expression.

@@

The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract.
The state of the contract is captured by stream semantics of AGREE.
When writing assumptions and guarantees, it is important to differentiate pre-state, before a contract updates its state, and post-state, after a contract updates its state, in response to the current input.
The \texttt{pre} naturally makes that distinction.
In general, assumptions regarding state should be evaluated in the pre-state of the component with the current inputs, and guarantees regarding state should be evaluated on the post-state of the component given the current inputs.
Guarantees should use \texttt{pre} anytime they need to reason about the current state relative to the previous state.

The assumption on \lineref{line:sw-assume-3} uses a followed-by expression so it is \texttt{true} at time 0 and then it is the truth value of the implication in all future instances.
The followed-by guards the \texttt{pre} as mentioned previously;
thus after time 0, the assumption depends on the presence of a request and the pre-state of the component.
The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}--\emph{not request since start}.

The guarantees on \linesref{line:sw-guarantee-1}{line:sw-guarantee-3} coincide events and assert well-formed output.
The guarantees on \linesref{line:sw-guarantee-4}{line:sw-guarantee-6} define temporal properties of the component.
\lineref{line:sw-guarantee-4} insists that a start happens with a request or one step after a request.
The guarantee uses the assumption on \lineref{line:sw-assume-3} and does not check for two requests in a row without a start as the assumption precludes that input behavior.
It differentiates with the followed-by what is required at time 0, the start must coincide with the request, with what is required after time 0, the start must coincide with the request or is a response to a request one step earlier.
The guarantee also does not force the start to always happen, it only says that if it does happen, it is under the defined conditions.

\lineref{line:sw-guarantee-5} forces the alert to sound if the start does not arrive within the one-step bound.
Together with \lineref{line:sw-guarantee-4} the contract model allows for non-alerting and alerting behavior.
\lineref{line:sw-guarantee-6} ensures if alert has ever happened, then it is happening in the present moment.
These six guarantees define the behavior of the SW component under the three assumptions and correspond to the informal descriptions given in \secref{sec:example}.
